### 2021-10-13 README.Rmd
initial commit

---
title: "README"
author: "Roland Hassan Nasser"
date: "13/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## digiRhythm

digiRhythm is an R library developed at Agroscope and provides a set of tools to 
analyze and visualize the rhythmic behavior of animals.

**TO DO:**

- [ ] Define data inside the library
- [ ] Create a function that read Icetag data stored in a CSV file.
- [ ] Add the previous function as an example to the README file.
- [ ] Create a function that visualize the actogram.
- [ ] Add the previous function as an example to the README file.
- [ ] Create test functions.
- [ ] Create a function that computes the degree of functional coupling.
- [ ] Create a function that computes the diurnality index.
- [ ] Shiny app?

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
#devtools::install_github("nasserdr/digiRhythm")
```
## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
#library(digiRhythm)
## basic example code
```