### 2023-05-25 jstatsoft/article.Rnw
Mid day results of changes by Hassan. Added a new amendment tracker with the date of today, fixed some references in the intro and improved the language for some parts.

\documentclass[article]{jss}

%% -- LaTeX packages and custom commands ---------------------------------------

%% recommended packages
\usepackage{orcidlink,thumbpdf,lmodern}
\usepackage[utf8]{inputenc}
\usepackage{algorithm}
\usepackage{algpseudocode}

%% another package (only for this demo article)
\usepackage{framed}
\usepackage{graphicx}
\graphicspath{ {./figures} }

%% new custom commands
\newcommand{\class}[1]{`\code{#1}'}
\newcommand{\fct}[1]{\code{#1()}}
% \newcommand{\ds}{'df516b_2'}
%% For Sweave-based articles about R packages:
%% need no \usepackage{Sweave}
\SweaveOpts{engine=R, eps=FALSE, keep.source = TRUE}
<<preliminaries, echo=FALSE, results=hide>>=
options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE)
library("digiRhythm")
@

%% -- Article metainformation (author, title, ...) -----------------------------

%% - \author{} with primary affiliation (and optionally ORCID link)
%% - \Plainauthor{} without affiliations
%% - Separate authors by \And or \AND (in \author) or by comma (in \Plainauthor).
%% - \AND starts a new line, \And does not.
\author{Hassan-Roland Nasser~\orcidlink{0000-0003-1821-3234}\\Agroscope - Digital Production
   \AND Marie Schneider~\orcidlink{0000-0003-4921-8353}\\Johann Heinrich von Th\"unen\\ Institut für \"Okologischen Landbau
   \And Marianne Cockburn~\orcidlink{0000-0003-3193-0641}\\Agroscope - Equine Research Group}
\Plainauthor{Hassan-Roland Nasser, Marie Schneider, Marianne Cockburn}

%% - \title{} in title case
%% - \Plaintitle{} without LaTeX markup (if any)
%% - \Shorttitle{} with LaTeX markup (if any), used as running title
\title{\pkg{digiRhythm}: an \proglang{R} package for rhythmicity assessment using the degree of functional coupling}
\Plaintitle{digiRhythm: an R package for rhythmicity assessment using the degree of functional coupling}
\Shorttitle{digiRhythm: an R package for rhythmicity assessment using the DFC}

%% - \Abstract{} almost as usual # comr: changed abstract throughout.
\Abstract{
Investigating an organism's rhythmicity yields valuable insights into its physiological and psychological state. What sets this measure apart is its applicability to various types of activity-related time series data, including accelerometers, GPS tracks, and smartwatches (e.g., step counts), among others. The Degree of Functional Coupling (DFC) represents a well-established method for assessing rhythmicity, encompassing several computational steps like frequency spectrum calculation and the determination of the statistical significance of spectral peaks. In this study, we introduce the \proglang{R} package \pkg{digiRhythm}, furnished with functions to compute rhythmicity from time series data. Complementary functions for actogram generation, computation and visualization of mean activities, and the calculation of a diurnality index, are also incorporated into this package. Tailored for seamless integration into the research workflow, \pkg{digiRhythm} facilitates the transformation of raw data into publication-ready content, thereby simplifying the application of these complex methods.}

%% - \Keywords{} with LaTeX markup, at least one required
%% - \Plainkeywords{} without LaTeX markup (if necessary)
%% - Should be comma-separated and in sentence case.
\Keywords{Lomb-Scargle periodogram, Degree of Functional Coupling, periodicity, circadian rhythm, harmonic frequencies, actogram, diurnality index, harmonic power}
\Plainkeywords{Lomb-Scargle periodogram, Degree of functional coupling, periodicity, circadian rhythm, harmonic frequencies, actogram, diurnality index, harmonic power}

%% - \Address{} of at least one author
%% - May contain multiple affiliations for each author
%%   (in extra lines, separated by \emph{and}\\).
%% - May contain multiple authors for the same affiliation
%%   (in the same first line, separated by a comma).

\Address{
  Hassan-Roland Nasser\\
  Agroscope, Digital Production\\
  Tanikon 1\\
  8352, Ettenhaussen \\
  Switzerland\\
  E-mail: \email{roland.nasser@agroscope.admin.ch, joa.stachowi@gmail.com}\\
  URL: \url{https://ch.linkedin.com/in/nasserha/, https://orcid.org/0000-0003-4921-8353}\\
  \newline

  Marie Schneider\\
  Johann Heinrich von Th\"unen Institut für \"Okologischen Landbau\\
  Trenthorst 32\\
  23847 Westerau\\
  Deutschland \\
  \emph{and}\\
  Universit\"at Hohenheim – Institut f\"ur Verfahrenstechnik der Tierhaltungssysteme \\
  Garbenstra{\ss}e 9 \\
  70599 Stuttgart\\
  Deutschland\\
  E-mail: \email{marie.schneider@thuenen.de}\\
  \newline

 Marianne Cockburn\\
  Agroscope, Equine Research Group \\
  Swiss National Stud \\
  Les Longs-Prés\\
  1580, Avenches \\
  Switzerland\\
  E-mail: \email{marianne.cockburn@agroscope.admin.ch}\\
  URL: \url{https://orcid.org/0000-0003-3193-0641}\\
  \newline

}

\begin{document}

\section[Introduction]{Introduction} \label{sec:intro}

Internal and external stimuli arouse adaptive behavioural patterns that can be monitored by analysing the rhythmicity of behaviours \citep{Berger2003}. Due to the increasing development of ICT (Information and Communication Technology) based on small, non-invasive sensors, large amounts of behavioural data are regularly recorded, and can be used to calculate parameters describing biological rhythms \citep{theurer2013remote,scheibe2006application, Berger2003}. \cite{sinz1976systemanalyse} developed a parameter able to quantify and investigate the rhythmicity of behavioural patterns, the Degree of Functional Coupling (DFC), which is based on telemetric measurements of activity and/or feeding of animals. The DFC provides an estimate of synchronization between an individual and its environment. It is calculated to each 24 h period within a timespan of a few days and predicts values between $0\%$ and 100 $\%$ \citep{scheibe1999comparative, Berger2003}. XXX suggested to use rolling seven day periods for this computation. An organism with a high DFC is highly synchronized with its environment, while lower DFC's indicate a lower synchronization. Breakdowns in the DFC response can be caused by stressors (e.g. handling, disease, etc.) that affect the time structure of behaviour \cite{Berger2003, scheibe1999comparative}. In fact, animals that live in stress-free conditions show activities all tuned to one zeitgeber (i. e. 24 hour periodicity), while their activity rhythm will deviate from the zeitgeber frequency if they are subjected to stressful situations \citep{scheibe1999comparative, Berger2003}.
A better understanding of changes in behaviour due to illness or distress is valuable to get further inisight in the organisms affective state % checke reference
(Grandin et al., 1989; Kilgore et al., 2005; Borderas et al., 2008; Robért et al., 2011). Bio-rhythmic analysis has the potential to provide the basis for the early detection of psychological and physiological disorders (Berger et al., 1999). With the increasing availability of sensor technologies to collect behavioral data continuously, further studies applying the DFC were performed \citep{sarout2018assessment, fuchs2022detecting}. These studies applied the DFC method in managed animal husbandry systems. However, the application of the DFC was always limited to those with extensive knowledge of statistical programming. An open source package specifically for this type of computation opens up the opportunities for livestock research to easily apply these algorithms. Therefore, we developed the \proglang{R} package \pkg{digiRhythm} \citep{dgm}, which was originally developed for livestock research, but has the potential to be applied in various domains.

This paper describes the functions of the package and is organized as follows. In section \ref{sec:dfc}, the mathematical details of computing the DFC are explained. Further, the statistical significance of frequency peaks needed for the DFC computation is explained. In section \ref{sec:usage}, a detailed overview of the \pkg{digiRhythm} package functions is provided. Finally, we indicate how this package can be useful for researchers in section \ref{sec:summary}.

\section{The degree of functional coupling}{\label{sec:dfc}}

The computation of the Degree of Functional Coupling (DFC) necessitates several steps. Initially, the Lomb-Scargle Periodogram \citep{lomb1976least, scargle1982studies, vanderplas2018understanding} of the activity is computed. Following this, the spectrum is analyzed, focusing on specific frequencies (including both harmonics and non-harmonics, with further details in Figure \ref{fig:reallsp}). Subsequently, the statistical significance of frequency peaks is calculated (\ref{sub:significance}). Ultimately, a ratio of the sum of frequency powers is determined (a detailed breakdown of this calculation is illustrated in Figure \ref{dfc_sum}). We commence by elucidating the ratio of frequency powers and then provide a detailed explanation on the computation of these frequency powers and their statistical significance.

As suggested by \cite{scheibe1999comparative}, the DFC can be computed using the following formula:
%
\begin{equation} \label{eq:dfc_for}
DFC = \frac{ssh}{sumsig}
\end{equation}
%
In parallel, another indicator termed 'harmonic power' can be calculated similarly:
%
\begin{equation} \label{eq:hp_for}
HP = \frac{ssh}{sumall}
\end{equation}
%
Here, $ssh$ represents the sum of powers of frequencies that are both harmonic and statistically significant, $sumsig$ refers to the sum of powers of significant frequencies (both harmonic and non-harmonic), and $sumall$ signifies the sum of the power of all frequencies in the spectrum.

Both DFC and HP are instrumental in studying the rhythmic behaviors of animals. The \pkg{digiRhythm} package facilitates the computation of both these indicators.

\subsection{The Lomb-Scargle Periodogram (LSP)}\label{sub:lsp}
The Lomb-Scargle Periodogram (LSP) marks the initial step in calculating the DFC and HP. It is a well-regarded tool within the astronomy community, commonly used for detecting and characterizing periodicity in unevenly sampled time-series data \citep{vanderplas2018understanding}.

In the time domain, we observe the magnitude of activity at specific time points (Figure \ref{fig:signalVslsp}, left). By applying the LSP, this activity can be transformed into the frequency domain (Figure \ref{fig:signalVslsp}, right). Here, the magnitude (power) of oscillations for specific frequencies present in the activity can be seen. For instance, a sine wave with a 24-hour period could be represented with a single-peak periodogram; this peak represents the frequency of a 24-hour sine wave. A similar principle can be illustrated for a sine wave with a 6-hour period, and so on. Moreover, activity composed of the summation of 24-hour and 6-hour sine waves will result in a periodogram with two peaks. Consequently, a periodogram allows us to infer the power of each possible sine wave present in the activity.

\begin{figure}
\includegraphics[width=1\columnwidth]{signalVslsp}
\caption{Exemplary harmonics (left) and their corresponding Lomb-Scargle Periodogram (LSP) (right). The first two rows depict regular sine waves activity signals with 24 hours and 6 hours periods, respectively, and their corresponding LSP. The bottom row illustrates the LSP (right) of two sine waves (of 24 hours period and 6 hours period) summed (left) together.}
\label{fig:signalVslsp}
\end{figure}

Activity signals from animals tend to be irregular, requiring conversion via the LSP to investigate periodic events within these signals. Figure \ref{fig:realsignal} illustrates an example of such data - the motion index (a proprietary metric of the overall animal activity measured in three dimensions \cite{icetagiowa}) of a cow, and its corresponding LSP is depicted in Figure \ref{fig:reallsp}.

\begin{figure}
\centering
<<realsignal, fig=TRUE, width=9, cache=TRUE>>==
data("df516b_2", package = 'digiRhythm')
df <- df516b_2
#We extract seven days from the dataset only for illustration purposes
sub_df <- df[1:672 , c(1,2)]
plot(sub_df$datetime, sub_df$Motion.Index, type = 'l',
xlab = 'DateTime', ylab = 'Motion Index')
@
\caption{\label{fig:realsignal} The \emph{Motion Index} activity column in the dataset \fct{df516b\_2}.}
\end{figure}

Mathematically, the LSP is computed using the following formula:

%
\begin{equation}{\label{eq:lsp}}
P_{LS}(f) = \frac{1}{2}
\Biggl\{
\frac{(\sum_n g_n \cos{(2 \pi nf[t_n - \tau])})^2}{\sum_n g_n \cos^2{(2 \pi nf[t_n - \tau])}}
+
\frac{(\sum_n g_n \sin{(2 \pi nf[t_n - \tau])})^2}{\sum_n g_n \sin^2{(2 \pi nf[t_n - \tau])}}
\Biggl\}
\end{equation}

Where $g_n$ is the $n_{th}$ sample in the activity, $t_n$ the $n_{th}$ time index, $f$ is the frequency (in hertz) and $\tau$, which depends on $f$ and ensures time-shift invariance, is presented by:

\begin{equation}{\label{eq:tau_lsp}}
\tau = \frac{1}{4\pi f} \tan^{-1} \frac{\sum_n \sin{(4\pi ft_n)}}{\sum_n \cos{(4\pi ft_n)}}
\end{equation}

Before performing the LSP, we must define the frequency grid to which we will apply formula \ref{eq:lsp}. This frequency grid ranges from 0 to $f_{max}$, the maximum frequency power that can be computed from the activity $g_n$. According to the Shannon-Nyquist theorem, $f_{max}$ is determined \citep{shannon1928certain, shannon1949mathematical}.

Figure \ref{fig:reallsp} represents the Lomb-Scargle periodogram computed from the activity shown in Figure \ref{fig:realsignal}.

\begin{figure}
\centering
<<reallsp, fig=TRUE, width=9, cache=TRUE>>==
lsp <- lomb_scargle_periodogram(df[1:672 , c(1,2)],
                         alpha = 0.01, sampling = 15, plot = TRUE)
head(lsp$lsp_data)
@
\caption{The Lomb-Scargle Periodogram of the activity data shown in Figure \ref{fig:realsignal}. Red bars show the amplitude of harmonic frequencies (with corresponding harmonic periods), while green bars represent non-harmonic frequencies. The dashed line indicates the significance threshold; frequencies with power below this threshold are considered non-significant. In this particular example, only two frequencies with significant power are evident; one of them - corresponding to a 12-hour period - is harmonic.}\label{fig:reallsp}
\end{figure}

\subsection{Understanding Harmonic Frequencies and the Underlying Intuition}\label{sub:harmonic}

Various animal species are characterized by circadian rhythms, signifying their behaviors align with a 24-hour cycle. This rhythmicity is manifested as power within the harmonic frequencies of the primary cycle (24 hours). Importantly, these harmonics refer to the frequencies linked to periods of 24 hours divided by integers less than 24 (e.g., 24/1, 24/2, …). At this juncture, it is crucial to revisit the concept of frequency power subsets that have been discussed earlier. These subsets are graphically represented in Figure \ref{dfc_sum}.

\begin{figure}[t]
\includegraphics[width=1\columnwidth]{sum_dfc}
\caption{Significant and non-significant frequency peaks depicted in a Lomb-Scargle Periodogram, contextualizing the rhythmicity in animal behaviors. The total frequency space includes frequencies with significant and non-significant peaks, the combined powers of which amount to $sumall$. The cumulative power of frequencies with significant peaks equates to $sumsig$. A subset of these significant peaks corresponds to harmonic frequencies (24h, 24h/2, 24h/3, \dots), and their summed power is represented by $ssh$.}
\label{dfc_sum}
\end{figure}

The connection between rhythmicity and frequencies should now be clear to the reader. If the frequencies deemed significant are exclusively harmonic (i.e., $ssh = sumsig$), the DFC equals 1, indicative of strong rhythmic behavior in the animal. Conversely, if the significant frequencies are not purely harmonic, the DFC value is less than 1, suggesting relatively low rhythmic behavior. This phenomenon could be indicative of sub-optimal physiological states.

\subsection{Evaluating the Statistical Significance of Frequency Peaks}\label{sub:significance}

The computation of the Degree of Functional Coupling (DFC) involves two critical steps. The first is the calculation of the periodogram, and the second is determining the statistical significance of the frequency peaks. A periodogram typically displays a distinct peak for each frequency, making the identification of valid activity components from noise components a fundamental task. This distinction becomes particularly challenging when dealing with a small sample size and a low signal-to-noise ratio, where noise peaks may mimic the magnitude of activity peaks. Consequently, it is imperative to utilize tools that can accurately measure the statistical significance of a peak.

Commonly, the False Alarm Probability (FAP) is employed to serve this purpose. FAP quantifies the likelihood that data devoid of any activity would generate a peak of comparable magnitude.

As reported by Scargle \cite{scargle1982studies}, the values of the un-normalized periodogram follow a $\chi^2$ distribution with two degrees of freedom. Therefore, the probability of observing periodogram values less than a certain $Z=P(f_0)$ in a data series constituted exclusively by Gaussian noise is expressed as:

\begin{equation}
P_{single} = 1 - \exp{-Z}
\end{equation}

Multiple methodologies can be used to calculate the FAP. Among these, the Baluev method, according to comparative studies by \citep{suveges2015comparative}, stands as one of the most effective and hence, has been implemented in the \pkg{digiRhythm} package.

As per Baluev \citep{baluev2008assessing}, the FAP can be formalized as follows:

\begin{equation}
  FAP(z) \approx 1 - [P_{single}(z)]\exp(-\tau(z))
\end{equation}

Where:
\begin{equation}
\tau(z) = W(1-z)^{(N-4)/2}\sqrt{z}
\end{equation}
and $W = f_{max}\sqrt{4\pi var(t)}$

\section{Functionalities and Usage of digiRhythm}\label{sec:usage}

The \pkg{digiRhythm} package offers a range of functionalities beyond merely calculating the Degree of Functional Coupling (DFC) and Harmonic Part (HP). In collaboration with animal behavior scientists, we have incorporated a variety of tools typically required for rhythmicity analysis. At present, the \pkg{digiRhythm} library includes the following features:

\begin{itemize}
\item \fct{lomb\_scargle\_periodogram}: This function not only computes the Lomb Scargle Periodogram (LSP) but also provides an intricate visualization of it.

\item \fct{dfc}: This function allows for both computation and graphical representation of the Degree of Functional Coupling and its harmonic part.

\item \fct{daily\_average\_activity}: This feature facilitates the calculation and plotting of the daily average activity.

\item \fct{diurnality}: With this function, users can compute and graphically present the diurnality index.

\item \fct{actogram}: This tool is designed for actogram visualization.
\end{itemize}

The \fct{dfc} function acts as a wrapper for the \fct{lomb\_scargle\_periodogram}, providing users with the opportunity to delve further into the Lomb-Scargle Periodogram, should they wish to explore it in greater detail.

\subsection{Format of Input Data}

As referenced in Section \ref{sec:intro}, the \pkg{digiRhythm} package necessitates that input data frames adhere to a specific structure. An example of a typically structured data frame, sampled at 15-minute intervals, is displayed below. This representative data frame is bundled with the \pkg{digiRhythm} package and can be retrieved using:

<<input_data, cache=TRUE>>==
data("df516b_2", package = 'digiRhythm')
df <- df516b_2
head(df)
str(df)
@

This excerpt presents a sample activity dataframe. The columns represent date-time, the motion index, and the number of steps measured during each 15-minute time interval.

To confirm that the data adheres to the required format, we've included a utility function capable of verifying its compatibility with \pkg{digiRhythm}. This check can be performed using the \fct{is\_dgm\_friendly} function, as demonstrated below:

<<is_dgm, cache=TRUE>>==
is_dgm_friendly(data = df, verbose = TRUE)
@

This function takes a data frame as an argument and verifies its \pkg{digiRhythm} compatibility, returning a boolean. The data is considered \pkg{digiRhythm}-friendly if it satisfies the following conditions:

\begin{itemize}
\item The data frame is not a NULL object.
\item The first column adheres to the POSIX standard.
\item The data frame contains at least two columns.
\item All columns, excluding the first one, are numeric.
\end{itemize}

The verbose argument aids in understanding why the data is or isn't considered \pkg{digiRhythm}-friendly. If verbose = TRUE, the function will provide information about each condition checked.

A warning message will be generated if the data encompasses less than seven days, since at least a week's worth of data is needed to compute the DFC and HP. However, other functionalities may still be executable.

\subsection[The dfc function]{The \fct{dfc} function}{\label{sub:dgm_fct}}
The DFC and HP are both implemented (\ref{alg:dfc}) in the \fct{dfc} function. The function evaluates a \pkg{digiRhythm} friendly data frame, by analyzing the start, and end date, and the variable for which the user wants to run the analysis. It then applies a Lomb Scargle Periodogram over a sliding window of seven days at a time. For instance, the DFC and HP are computed for the days one to seven, then for the days two to eight, and so on. The last six days of the data set would not be considered because it is not recommended to use the LSP (in the context of computing the DFC) for less than seven days. Figure \ref{fig:dfc} shows a typical DFC/HP output.

\begin{algorithm}
\caption{The Degree of Functional Coupling algorithm}\label{alg:dfc}
\begin{algorithmic}
\Require {data}
\Ensure {data is compatible with the library}
\State $days \gets Total\ Number\ of\ Unique\ Days$
\State $output \gets NULL$
\For{\texttt{day in 7:days}}
        \State \texttt{Select data from the day ($day-7$) to day ($day$)}
        \State \texttt{Compute the Lomb Scargle Periodogram of the selected data}
        \State \texttt{Compute the sum of power for all frequencies $sumall$}
        \State \texttt{Select Harmonic frequencies and compute $ssh$}
        \State \texttt{Select Harmonic frequencies and compute $sumsig$}
        \State {dfc $\gets$ ssh/sumsig}
        \State {hp $\gets$ ssh/sumall}
        \State {Output[day] $\gets$ \{day, dfc, hp\}} \Comment{Append the data frame  with a new line}
      \EndFor
\end{algorithmic}
\end{algorithm}

\subsection[Using digiRhythm to compute DFC and HP]{Using \pkg{digiRhythm} to compute DFC and HP}{\label{sub:dgm_dfc_hp}}

The DFC and HP can be computed using the \fct{dfc} function. A recommended workflow is shown below. After reading the data and verfying its compatibility with the library, the data (originally sampled with 1 min) is resampled to 15 minutes. In this context, resampling means simple summation. For instance, if the dataset contains the number of steps with 1 minute sampling period, resampling to 15 minutes means adding the steps during each 15 minutes interval. Further, an activity is chosen (in this example, we choose the activity from the second column (motion index)), then the DFC is computed using the \fct{dfc} function as shown below:

<<dfc, cache=TRUE>>==
<<dfc_snippet, results=hide>>=
df_resampled <- resample_dgm(df, 15)
activity = names(df_resampled)[2]
my_dfc <- dfc(df_resampled, activity = activity,
              plot = TRUE, verbose = FALSE)
@

The returned data frame  contains the DFC and HP measured along a sliding window of seven days as follows:

<<print_dfc, cache=TRUE>>==
head(my_dfc$data)
print(class(my_dfc))
@

The returned variable \code{my_dfc} contains both the data and the plot as it is a \code{ggplot2} object.

\begin{figure}[t!]
\centering
<<plot_dfc, fig=TRUE, cache = TRUE, width=9, cache=TRUE>>==
library("ggplot2")
plot(my_dfc)
@
\caption{\label{fig:dfc} The DFC and HP of the dataset \fct{df516b\_2}.}
\end{figure}

It is possible to access the computed data and the user can adjust the plot features.
Calling the data field from the variable \code{my_dfc} will allow accessing the computed DFC and HP as shown below:

It is, therefore, easy to access the resulting DFC and HP and use them in further analysis like statistical modeling or more customized workflows. Because the returned object is a \pkg{ggplot2} \citep{wickham2010layered}, it is also possible to change the plot directly by adding a layer on the aesthetics of the \code{my_dfc} object as shown below:

\begin{figure}[h!]
\centering
<<dfc_plot_modified, fig=TRUE, width=9, cache=TRUE>>==
new_plot <- my_dfc +
   theme(
     text = element_text(family = "Times", size = 8),
     axis.text.y = element_text(size = 15, colour = "red"))
print(new_plot)
@

\caption{\label{fig:dfc_modified} Output plot from Figure \ref{fig:dfc} modified by adding a layer of aesthetics causing some modifications to the \fct{ggplot} output returned by the \fct{dfc}. In particular, the font family was changed to \emph{Times}, the font size to 8 points, and the y-axis text was changed to a bigger size (15) and colored red.}
\end{figure}

This is an illustrative yet straightforward example to sketch the feature of reusing the graphical outputs of the library, but the application is useful when users need to produce publication-quality pictures in a specific format. It is important to mention that this feature is possible because the functions were designed in a way they return \fct{ggplot} objects rather than base \proglang{R} objects.

\subsection[Recommendations to use the dfc function]{Recommendations to use the dfc function}{\label{sub:recom_dfc}}

Although the \fct{dfc} function can accept data with any sampling frequency, it is suggested by several authors \citep{sinz1976systemanalyse, sarout2018assessment, fuchs2022detecting, scheibe1999comparative} that optimal insights could be obtained using seven days sliding window with 15 minutes sampling period. For the sake of simplicity, an option for changing the sliding window length (for example, 6 or 8, or more days) was not implemented. However, this could be done in future versions if proven useful.

\subsection[Actogram visualization]{Actogram visualization}{\label{sub:acto}}

The actogram is often the first-to-go tool when working with animal activity. It is a tiled 2D representation where the x-axis represents the time of the day, the y-axis represents the number of days (or date), and the tiles are color-coded according to the activity intensity at each sample.

Figure \ref{fig:actogram} shows a code snippet for visualizing the actogram of the dataset \fct{df516b\_2} and its corresponding actogram.

\begin{figure}[h!]
\centering
<<actogram, fig=TRUE, width=9, cache=TRUE>>==
start <- "2020-05-01" #year-month-day
end <- "2020-06-14" #year-month-day -->
activity_alias <- 'Motion Index'
actogram(df, activity, activity_alias , start, end, save = NULL)
@
%
\caption{\label{fig:actogram} The actogram of the of the dataset \fct{df516b\_2}.}
\end{figure}

\subsection[Average activity]{Average activity}{\label{sub:avg}}

The average activity is an indicator of how active an animal is on average during times of the day. The average activity is computed using the following formula:

\begin{equation}
  activity_{average}(T_i) = \frac{1}{d} \sum_d {A(T_{id})}
\end{equation}

Where $T_i$ is a specific time of the day, $d$ is the number of days, and $A(T_{id})$ is the activity value at the time $T_i$ during the day $d$.

The function \fct{daily\_average\_activity} computes and plots this indicator as shown in Figure \ref{fig:avg_act}.

\begin{figure}[h!]
\centering
<<avg_activity, fig=TRUE, width=9, cache=TRUE>>==
daily_average_activity(df, activity, activity_alias,
                                 start, end, save = NULL)
@
\caption{\label{fig:avg_act} The average activity of the dataset \fct{df516b\_2}.}
\end{figure}

\subsection[Diurnality Index (DI)]{Diurnality Index (DI)}{\label{sub:di}}

The Diurnality Index is also an important measure when it comes to assessing animals' rhythmicity. This measure indicates whether the animal exercises its activity more during diurnal (daytime) or nocturnal (nighttime) periods. Using the Diurnality function, the user can specify day and night time. The user can also omit some periods during the day, such as milking time, where the milking process controls the animal's activity rather than its behavior. The Diurnality index is computed according to \cite{hoogenboom1984seasonal}:

\begin{equation}
  DI = \frac{\frac{C_d}{Td} - \frac{C_n}{Tn}}{\frac{C_d}{Td} + \frac{C_n}{Tn}}
\end{equation}

Where $C_d$ and $C_n$ are the summations of activity during the daytime and nighttime, respectively. $T_d$ and $T_n$ are the numbers of activity samples during the daytime and nighttime. The number of samples is usually measured from the sampling time. For instance, if the activity is measured with 15 min time sample, and the daytime is considered to be between 06h00 AM and 04h00 PM, then $T_d$ is considered equal to 4 samples per hour × 10 hours = 40 samples.
The Diurnality function takes the daytime and nighttime as arguments, with default daytime as 06h30 AM till 04h30 PM and default nighttime as 18h00 pm till 05 AM the next day. To change these times, the configuration should follow an HH:MM:SS format as shown below:

<<setting_times, cache=TRUE>>==
day_time <- c("06:30:00", "16:30:00")
night_time <- c("18:00:00", "05:00:00")
@

Figure \ref{fig:di} shows how to compute and visualize the diurnality index for an example dataset.

\begin{figure}[h!]
\centering
<<diurnality, fig=TRUE, cache=TRUE>>==
my_di <- diurnality(df, activity, day_time, night_time)
@
%
\caption{\label{fig:di} The Diurnality index of the dataset \fct{df516b\_2}.}
\end{figure}

\subsection[sliding Diurnality Index (DI)]{sliding Diurnality Index (DI)}{\label{sub:sliding_di}}

The function \fct{sliding\_DI} computes the diurnality index as described above. However, instead of using a single definition of day and night for the complete computation period, a data frame defining different definitions of the start/end of day/night is included. The data frame must include five columns, one date column, and four columns of timestamps for start/end of day/night. All of these columns must be formatted as POSIXct.

This function, which includes different definitions, is useful in various settings, as it allows to improve the computation accuracy. For example, in Europe, it could be used to reduce the bias emerging by converting summertime and wintertime to a time zone, such as GMT, or by defining sunrise and sunset.

%% -- Summary/conclusions/discussion -------------------------------------------

\section{Discussion} \label{sec:summary}
The \pkg{digiRhythm} package can make a significant contribution to the assessment of rhythmicity, as it makes the analysis of rhythmicity more accessible to researchers. The tool is easy to use, especially because it is delivered as an \proglang{R} package. The theoretical details and intuition behind the DFC is explained in section \ref{sec:dfc}. The usage of the package is explained in the section \ref{sec:usage}. Further examples can be found in the library's vignettes.

The package enables the calculation of rhythmicity from activity data. To date, this is purely a computation of activity patterns that have been correlated with physiological factors in previous studies. This indicator can be calculated from any type of regularly measured time series data, if they are brought in a suitable format, which can be assessed with the  \fct{is\_dgm\_friendly} function. However, we have no evidence if these activity patterns are intrinsically (such as going to sleep) or extrinsically (such as waking up to an alarm) motivated. Such extrinsic factors could affect rhythmic activity patterns and blur physiological rhythmicity patterns. To date, the package provides the option to exclude extrinsically structured time spans from the analysis. However, in order to increase the understanding of rhythmicity  future studies should investigate how these extrinsic factors affect rhythmicity calculations.

The required frequency of data collection is of utmost importance in sensor technologies. Sensors are required to be small and further expected to run without any sensor based maintenance, such as data reading or charging of the device. This in turn means that data collection frequency should be low. In the current study an extremely low data collection frequency is suggested (15 minutes) future studies should evaluate how low collection frequency can be set to create reliable rhythmicity calculations. Further it would be interesting to evaluate if activity data collected from different sources, such as  access of feeding stations or coffee maschines, compute similar rhythmicity outputs in the same organism.

Seven days of continuous data collection are required to compute one DFC value. This implies that very long data collection intervals are required to detect rhythmic patterns. Reducing the number of days required to compute one DFC value would make this rhythmicity computation much more accessible in practice an research. The impact of reducing the number of days required to compute the DFC should be investigated in future research.

Other metrics for calculating circadian rhythms have been previously published. For example, Interdaily Stability (IS) is mainly used in human research \cite{witting1990alterations}. The Fourier-Based Approximation with Thresholding (FBAT) was designed to calculate the circadian rhythms of dairy cows \cite{wagner2021detection}. Their database depends on the 24h day, but none of them uses the 24h within the calculation. The calculation of harmonic frequencies, which sets the activity signal in relation to the external 24h day is unique. Therefore, it would be interesting to investigate the correlation between the DFC and rhythmicity computed on data from various computation techniques. For instance, \cite{moller2022diurnal} utilized a range of techniques, including set logic, circular cross-correlation, circular clustering, functional enrichment analyses, and least squares regression, on transcriptomic time series data from humans and baboons. Additionally, GPS data of free ranging animals \cite{plaza2022gps} or data of real time positioning systems in a barn \cite{wagner2021detection} could be used to estimate rhythmicity. Comparing the correlations between rhythmicity measures derived from different types of data and between various techniques could be a useful research area. In addition, further studies will be conducted, to validate the DFC calculation, by using simulated data.

The current package has been developed for farm animals, but the principles of rhythmicity apply for any organism, including humans. Therefore we highly recommend future research to evaluate the \pkg{digiRhythm} package in other species. Many humans nowadays continuously collect activity data with their smart-watches or the phones. Thus, it would be interesting to evaluate the application of the \pkg{digiRhythm} package in humans, offering an objective indicator of human welfare. A previous review highlights the potential of such rhythmicity computations in humans (https://journals.sagepub.com/doi/pdf/10.1177/0748730420940483).

At the time of writing this paper, the current version (v1.1) of the library does not provide the capability to compute rhythmicity using techniques such as cosinor analysis (\cite{refinetti2016}). However, there is potential for future expansion of the library's functionality to incorporate additional techniques. Researchers are encouraged to actively contribute by submitting new data, codes, or methods to the GitHub repository. These contributions will be reviewed by the package maintainers and, if approved, integrated into the \pkg{digiRhythm} package.

\bibliography{refs}
\newpage

\end{document}