### 2024-07-16 README.Rmd
Prettify all the codes with the package styler

---
title: "readme"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## digiRhythm

digiRhythm is an R library developed at Agroscope and provides a set of tools to 
analyze and visualize the rhythmic behavior of animals.

**TO DO CORE FUNCTIONALITIES:**

- [X] Create a function that read Icetag data stored in a CSV file.
- [X] Add the import_raw_icetage_data as an example to the README file.
- [X] Create a function that computes the degree of functional coupling.
- [X] Add the the dfc as an example to the README file.
- [x] Create a function that computes the diurnality index.
- [x] Add the the dfc as an example to the README file.
- [ ] Consider adding other sample data sets (check which ones appear to have
non-noisy data from the previous analysis and visualizations).
- [ ] Enhance the actogram (from the original codes).
- [ ] Add a function that fills NA values in the activity data set.
- [ ] Add a function that creates a histogram for the significant and harmonic
frequencies within the DFC (Patricia's case).

**TO DO - VISUALIZATION FUNCTIONS:**

- [X] Create a function that visualize the actogram.
- [ ] Create a function that visualize the actogram + average activity.
- [ ] Create a function that visualize the actogram + average activity + DFC.
- [ ] Create a function that visualizes all daily activities with control over
the number of columns.

**TO DO - UTILITIES AND DOCUMENTATIONS:**

- [X] Define data inside the library
- [X] Add and example about the import_raw_icetag_data to the README.
- [x] Create a function to test if a data set is digiRhythm friendly.
- [x] Create a proper documentation for the data set.
- [x] configure data sets with lazy loading
- [ ] Create test functions for all the functionalities.
- [ ] Add a note about the fact that visualizations were created in a way they
are aligned with the standards of common journals and give details about these 
settings + Add a note where the code can be customized to change these settings.
- [ ] Make a series of videos about the digiRhythm library.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

```{r}
# Uncomment the below two lines if you use the library for the first time of if
# you want to update the library
# install.packages("devtools")
# devtools::install_github("nasserdr/digiRhythm", dependencies = TRUE)
```
## Example

This is a basic example which shows you how to solve a common problem:

```{r example, message=FALSE, warning=FALSE, paged.print=FALSE}
library(digiRhythm)

# The file name with the path
url <- "https://github.com/nasserdr/digiRhythm_sample_datasets/raw/main/516b_2.csv"
download.file(url, destfile = "516b_2.csv")
filename <- file.path(getwd(), "516b_2.csv")

# The columns that we are interested in
colstoread <- c("Date", "Time", "Motion Index", "Steps")

# Reading the activity data from the csv file
data <- import_raw_icetag_data(filename = filename, skipLines = , act.cols.names = colstoread, sampling = 15)

print(head(data))
```

This is an example on how to visualize the actogram

```{r visualize, message=FALSE, warning=FALSE, paged.print=FALSE}
data("df691b_1", package = "digiRhythm")
df <- remove_activity_outliers(df691b_1)
df_act_info(df)
activity <- names(df)[2]
start <- "2020-08-25"
end <- "2020-10-11"
save <- TRUE
outputdir <- "testresults"
outplotname <- "myplot"
width <- 10
device <- "tiff"
height <- 5
actogram(df, activity, start, end,
  save = FALSE,
  outputdir = "testresults", outplotname = "actoplot", width = 10,
  height = 5, device = "tiff"
)
```

This is an example on how to compute the degree of functional coupling.

```{r dfc_example, results = FALSE, warning=FALSE, paged.print=FALSE}
data("df516b_2", package = "digiRhythm")
df <- remove_activity_outliers(df516b_2)
df_act_info(df)
activity <- names(df)[2]
my_dfc <- dfc(df, activity, sampling = 15, show_lsp_plot = FALSE)

# You may want to explore the two list inside my_dfc.
# DFC and SPECTRUM are saved inside my_dfc, each as a list
```

This is an example on how to compute the diurnality index:

```{r dindex_example, results = FALSE, warning=FALSE, paged.print=FALSE}
data("df516b_2", package = "digiRhythm")
df <- remove_activity_outliers(df516b_2)
df_act_info(df)
activity <- names(df)[2]
d_index <- diurnality(data, activity, plot = TRUE)
```

This is an example on how you can resample your data:

```{r resample, results = FALSE, warning=FALSE, paged.print=FALSE}
data("df516b_2", package = "digiRhythm")
df <- df516b_2
df <- remove_activity_outliers(df)
new_sampling <- 30
new_dgm <- resample_dgm(df, new_sampling)
```