### 2024-07-25 README.Rmd
Updated the vignettes and the readme page

---
title: "readme"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## digiRhythm

digiRhythm is an R library developed at Agroscope and provides a set of tools to 
analyze and rhythmic behavior of animals using the Degree of Functional Coupling. 
It also provides a set of vizualization tools such as Actogram, Diurnality Index,
Average Daily Activity, and Daily Activity Wrap Plot.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

```{r}
# Uncomment the below two lines if you use the library for the first time of if
# you want to update the library
# install.packages("devtools")
# devtools::install_github("nasserdr/digiRhythm", dependencies = TRUE)
```

## Example

You can access few examples from the vignettes provided along with the package.

[Vignettes](https://cran.r-project.org/web/packages/digiRhythm/index.html)